Simulation of Congenital Heart Defect Corrective Surgeries Using Thin Shell Elements

### 2012-03-02 Doc/Miccai/simulation.tex
Improved simulation section.
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@382 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

In order to repair malformations of the great arteries surgeons perform several low-level procedures on the blood vessels. These include arbitrary incisions, resections of whole blood vessel parts, attaching patches and reconnection of blood vessels end-to-end or end-to-side. From a technical view, all cutting operations can be done prior to simulation by splitting shell elements along their edges which correlates to the actual incision. New shell elements can be added to a mesh or rearranged to align their edges along an incision. On the contrary, joining and suturing of blood vessels and patches typically deform blood vessel walls to a certain amount which has to be simulated. Recent simulation systems \cite{} realize suturing through attachment of contracting springs that have no physical correlation. Springs introduce surplus energy to a physical system and special care has to be taken during simulation to compensate for that.

We present a new approach that overcomes these difficulties by connecting edges of shell elements along two opposite edges topologically (see figure \ref{fig-}. Since there is a distance between two joining edges in general, a topologically connection alters the shape and size of affected shell elements. Note that in this simulation state there is temporarily no physical correlation. We then start a controlled relaxation process in which the shell elements progressivly recover to their rest shape so that there is no strong increase in internal forces and the simulation remains stable. This is achieved by linear interpolation between the strongly deformed shell elements and their original rest shape. The steady state of the simulation corresponds to an energetical equilibrium of the shell element deformation and the simulation is finished. The effect is similar to contracting springs but there is no such problem like surplus energy after the joining process is done.

>>>>

In order to repair malformations of the great arteries surgeons perform several low-level procedures on blood vessels. These include arbitrary incisions, resections of whole blood vessel parts, attaching patches and reconnection of blood vessels end-to-end or end-to-side. From a technical view, all cutting operations can be done prior to simulation by splitting shell elements along their edges which correlates to the actual incision. New shell elements can be added to a mesh or rearranged to align their edges along an incision. On the contrary, joining and suturing of blood vessels and patches typically deform blood vessel walls to a certain amount which has to be simulated. Recent simulation systems \cite{Sorensen2006,Mosegaard2004,Li2009} realize suturing through attachment of contracting springs that have no physical correlation. Springs introduce surplus energy to a physical system and special care has to be taken during simulation to compensate for that.

We present a new approach that overcomes these difficulties by performing topological changes on the mesh to connect shell elements along joining edges (see figure \ref{fig-}. Since there is a certain distance between two joining edges in general, a topological modification alters the shape and size of affected shell elements. Note that in this simulation state there is temporarily no correlation to reality. We then start a controlled relaxation process during which the shell elements progressivly converge to their energetical equilibrium. To keep the simulation stable we avoid a strong increase in internal forces by linearly interpolate the rest shape of the connected mesh to the initial rest shape before joining. The effect is similar to contracting springs but there is no such problem like surplus energy after the joining process is done.
